---
title: "Assignment 2"
author: "Leona Ma"
date: "10/7/2021"
output: 
    html_document:
      toc: yes 
      toc_float: yes
      keep_md : yes 
    github_document:
      html_preview: false
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Obejectives: conduct data wrangling and visualize the data with key questions in mind.

# Part 1 Data Wrangling

## Step1 reading and merging the data sets
```{r reading and merging}
library(data.table)
library(dplyr)
library(dtplyr)

chs_ind <- data.table::fread("chs_individual.csv")
chs_reg <- data.table::fread("chs_regional.csv")

chs <- merge(chs_ind, chs_reg, by="townname")
```
The number of observations in chs and chs_ind is the same (1200), so there is no repeat data.

## Step 2 creating new categorical varible "obesity_level and summarize"
```{r creating var1}
chs$obesity_level <- as.factor(ifelse(chs$bmi < 14, 'underweight',
                               ifelse(chs$bmi < 22, 'normal', 
                               ifelse(chs$bmi < 24, 'overweight','obese' ))))

summary <- chs[,.(
  minimum = min(bmi, na.rm=TRUE),
  maximum = max(bmi, na.rm=TRUE),
  total = length(bmi)), by=obesity_level]
summary
                            
```

## Step 3 Create another categorical variable "smoke_gas_exposure"
```{r creating var2}
chs[, smoke_gas_exposure := fifelse(smoke == 0 & gasstove == 0, "Neither",
        fifelse(smoke == 0 & gasstove == 1, "gas_only",
               fifelse(smoke == 1 & gasstove == 0, "smoke_only",
                       fifelse(smoke == 1 & gasstove == 1, "both", NA_character_))))]
#summarizing 
chs[,.(total = length(smoke)), by=smoke_gas_exposure]
```
There are four categories and some missing values. 

## Step 4 Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
```{r townname}
chs[, .(
    fev_avg_townname      = mean(fev, na.rm=TRUE),
    fev_sd_townname       = sd(fev, na.rm=TRUE),
    asthma_avg_townname   = mean(asthma, na.rm=TRUE),
    asthma_sd_townname    = sd(asthma, na.rm=TRUE)),
    by = townname
    ] 
```
```{r sex}
chs[, .(
    fev_avg_sex      = mean(fev, na.rm=TRUE),
    fev_sd_sex       = sd(fev, na.rm=TRUE),
    asthma_avg_sex   = mean(asthma, na.rm=TRUE),
    asthma_sd_sex    = sd(asthma, na.rm=TRUE)),
    by = male
    ] 
```

```{r obesity}
chs[, .(
    fev_avg_obesity      = mean(fev, na.rm=TRUE),
    fev_sd_obesity       = sd(fev, na.rm=TRUE),
    asthma_avg_obesity   = mean(asthma, na.rm=TRUE),
    asthma_sd_obesity    = sd(asthma, na.rm=TRUE)),
    by = obesity_level
   ]
```
```{r smoke gas exposure}
chs[, .(
    fev_avg_smoke_gas_exposure      = mean(fev, na.rm=TRUE),
    fev_sd_smoke_gas_exposure       = sd(fev, na.rm=TRUE),
    asthma_avg_smoke_gas_exposure   = mean(asthma, na.rm=TRUE),
    asthma_sd_smoke_gas_exposure    = sd(asthma, na.rm=TRUE)),
    by = smoke_gas_exposure
   ]
```

# Part 2 Looking at the Data (EDA)

```{r}
dim(chs)
head(chs)
tail(chs)
str(chs)
summary(chs$obesity_level)
summary(chs$fev)
summary(chs$smoke_gas_exposure)
summary(chs$pm25_mass)
```

## Step1 Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r}
library(tidyverse)
ggplot(data=chs, mapping=aes(x = bmi, y = fev)) +
  geom_point() + 
  facet_wrap(~townname) +
  geom_smooth(method='lm', formula = y~x)+
  labs(title = "scatterplots of BMI vs FEV by townname", x="BMI", y="FEV")
```

From those figures we can see that BMI has a relatively positive association with FEV in all of these towns.

## Step2 Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}
ggplot(data=chs) +
  geom_histogram(aes(x=fev, fill=obesity_level), binwidth = 100)+
  scale_fill_brewer(palette = "Spectral")
  labs(title = "Stacked histograms of FEV by BMI", x="FEV")
```

From this stacked histogram we can see that people who are underweight has a relatively low fev, and people who are overweight and obese has a relatively high fev. People with normal bmi has fev that concentrated on normal range. 

```{r}
ggplot(data=chs[-which(is.na(chs$smoke_gas_exposure)=="TRUE"),]) +
  geom_histogram(aes(x=fev, fill=smoke_gas_exposure), binwidth = 100)+
  scale_fill_brewer(palette = "GnBul")+
  labs(title = "Stacked histograms of FEV by smoke/gas exposure", x="FEV")
```

This figure does not show different distributions among the categories. They all seem normally distributed. 

## Step3 Barchart of BMI by smoke/gas exposure.
```{r}
ggplot(data=chs[-which(is.na(chs$obesity_level)=="TRUE"),]) +
  geom_bar(aes(x=obesity_level, fill=smoke_gas_exposure))+
  scale_fill_brewer(palette = "Spectral")+
  labs(title = "Barchart of BMI by smoke/gas exposure", x="BMI")
  
```

This plot shows that the distribution of smoke/gas exposure is not different among the BMI categories. 

## Step4 Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}
ggplot(data=chs[-which(is.na(chs$obesity_level)=="TRUE"),],mapping = aes(x = obesity_level, y = fev))+
  geom_boxplot()+
  labs(title = "statistical summary graph of FEV by BMI")
```

This figure gives similar information as the stacked histogram that people who are underweight has a relatively low fev, and people who are overweight and obese has a relatively high fev. People with normal bmi has fev that concentrated on normal range. 

```{r}
ggplot(data=chs[-which(is.na(chs$smoke_gas_exposure)=="TRUE"),],mapping = aes(x = smoke_gas_exposure, y = fev))+
  geom_boxplot()+
  labs(title = "statistical summary graph of FEV by smoke/gas exposure category")
```

This figure does not show different distributions among the categories.

## Step5 A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r}
library(leaflet)

tem.pal <- colorNumeric(c('green','yellow','red'), domain=chs$pm25_mass)
leaflet(chs) %>% 
  addTiles() %>%
  addCircles (lat = ~lat, lng = ~lon, color = ~tem.pal(pm25_mass),
               label = ~pm25_mass, 
               opacity=1, fillOpacity = 1, radius = 5000) %>%
  addLegend('bottomleft', pal= tem.pal, values = chs$pm25_mass,
             title ='concentrations of PM2.5 mass of the CHS communities', opacity=0.5)
```
The mass of PM2.5 is concentrated around Riverside and San Bernardino region, and spread out to the sides. 

## Step6 Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r}
ggplot(data=chs, mapping=aes(x=pm25_mass, y=fev))+
  geom_point()+
  geom_smooth(method='lm', formula = y~x)
```

We can see from the figure that PM2.5 mass does not have obvious association with FEV.
